Cluster Analysis: Identifying Parkinson's Disease Subtypes

### 2015-06-26 writeup/parkinsons.tex
Add hierarchical clustering, change ht to h
LaTeX warns for h but I feel like there's no problem...it's just an
indication that the figure must be placed on the [t]op of the next page

Right?

\begin{table}[ht]
  \centering
  \begin{tabular}{l|l|l|l}
    Name & Type & Format & Description \\
    \hline
    nms\_d1 & byte & \%8.0g & cardiovascular \\
    nms\_d2 & byte & \%8.0g & sleep/fatigue \\
    nms\_d3 & byte & \%8.0g & mood/cognition \\
    nms\_d4 & byte & \%8.0g & percep/hallucinations \\
    nms\_d5 & byte & \%8.0g & attention/memory \\
    nms\_d6 & byte & \%8.0g & gastrointestinal \\
    nms\_d7 & byte & \%8.0g & urinary \\
    nms\_d8 & byte & \%8.0g & sexual function \\
    nms\_d9 & byte & \%8.0g & miscellaneous \\
    tremor & float & \%9.0g & tremor \\
    bradykin & float & \%9.0g & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & \%9.0g & rigidity \\
    axial & float & \%9.0g & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
    pigd & float & \%9.0g & postural instability and gait difficulty \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[ht]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.76&  3.32&   0-24 \\
nms\_d2&   8.71&  8.76&   0-48 \\
nms\_d3&   8.70& 11.83&   0-60 \\
nms\_d4&   1.65&  3.94&   0-33 \\
nms\_d5&   5.22&  7.44&   0-36 \\
nms\_d6&   5.67&  6.92&   0-36 \\
nms\_d7&   8.02&  9.09&   0-36 \\
nms\_d8&   3.57&  5.97&   0-24 \\
nms\_d9&   6.99&  7.74&   0-48 \\
tremor&   2.59&  2.63&   0-12 \\
bradykin& 2.49&  1.39&   0-6 \\
rigidity& 2.34&  1.36&   0-6 \\
axial&    3.28&  2.75&   0-12 \\
pigd&     3.36&  2.77&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{pca-eigenvalues.png}
  \caption{Scree test: eigenvalues by factor}
  \label{fig:pca-eigenvalues}
\end{figure}

@@

\section{$k$-means}
\subsection{Identifying optimal number of clusters}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-wss-error.png}
  \caption{Scree test: WSS error by cluster size}
  \label{fig:kmeans-wss-error}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.png}
  \caption{Gap statistic by cluster size}
  \label{fig:gap-statistic}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{asw.png}
  \caption{Average silhouette width by cluster size}
  \label{fig:asw}
\end{figure}

@@

\subsection{Cluster statistics}
% CLUSTERS: 2
% ================================
% Sizes: 201 580
% WithinSS: 4248.585 4132.434
% Sum WithinSS: 8381.019
% CLUSTERS: 3
% ================================
% Sizes: 420 231 130
% WithinSS: 2618.368 1973.82 3076.542
% Sum WithinSS: 7668.73
% CLUSTERS: 4
% ================================
% Sizes: 61 372 145 203
% WithinSS: 1481.25 1845.389 2147.988 1609.555
% Sum WithinSS: 7084.183
\begin{table}[ht]
  \centering
  \begin{tabular}{l|l|l|l}
    $k$ & $n$ & Within SS & sum(Within SS) \\
    \hline
    2 & 201/580 & 4248.585/4132.434 & 8381.019 \\
    3 & 420/231/130 & 2618.368/1973.82/3076.542 &  7668.73 \\
    4 & 61/372/145/203 & 1481.25/1845.389/2147.988/1609.555 & 7084.183
  \end{tabular}
  \caption{Cluster statistics}
  \label{tab:cluster-statistics}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-2.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 2$}
  \label{fig:kmeans-silhouette-2}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-3.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 3$}
  \label{fig:kmeans-silhouette-3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-4.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 4$}
  \label{fig:kmeans-silhouette-4}
\end{figure}

\subsection{Decision trees based on clusters}
% seed = 911
% CLUSTERS: 2
% ================================
% Complexity Parameter: 0.03482587
% 10-fold CV error: 0.134443
% Root node error: 0.2573624
% CLUSTERS: 3
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.1946223
% Root node error: 0.4622279
% CLUSTERS: 4
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.2483995
% Root node error: 0.5236876
\begin{table}[ht]
  \centering
  \begin{tabular}{l|l|l|l|l|l}
    $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    validation} & Root Feature &
    Root Error & Figure \\
    \hline
    2 & 0.0348 & 0.134 & axial $\geq$ 0.44 & 0.257 & Figure~\ref{fig:kmeans-dtree-2} \\
    3 & 0.0100 & 0.194 & bradykin $<$ 0.0041 & 0.462 & Figure~\ref{fig:kmeans-dtree-3} \\
    4 & 0.0100 & 0.248 & bradykin $<$ 0.0041 & 0.523 & Figure~\ref{fig:kmeans-dtree-4} \\
  \end{tabular}
  \caption{$k$-kmeans decision trees statistics}
  \label{tab:k-means-dtrees}
\end{table}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-2.pdf}
  \caption{Decision Tree from $k$-means clustering, 2 clusters}
  \label{fig:kmeans-dtree-2}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-3.pdf}
  \caption{Decision Tree from $k$-means clustering, 3 clusters}
  \label{fig:kmeans-dtree-3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-4.pdf}
  \caption{Decision Tree from $k$-means clustering, 4 clusters}
  \label{fig:kmeans-dtree-4}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-2.pdf}
  \caption{Cluster Summaries, $k = 2$}
  \label{fig:kmeans-summaries-2}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-3.pdf}
  \caption{Cluster Summaries, $k = 3$}
  \label{fig:kmeans-summaries-3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4.pdf}
  \caption{Cluster Summaries, $k = 4$}
  \label{fig:kmeans-summaries-4}
\end{figure}

\section{Affinity Propagation}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{ap-silhouette.pdf}
  \caption{AP silhouette plot, $k = 8$}
  \label{fig:ap-silhouette}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{l|l}
  Cluster & Size \\
  \hline
  1 & 54 \\
  2 & 68 \\
  3 & 25 \\
  4 & 259 \\
  5 & 102 \\
  6 & 68 \\
  7 & 185 \\
  8 & 20 \\
  \end{tabular}
  \caption{AP Cluster Sizes}
  \label{tab:ap-cluster-sizes}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{ap-summaries.pdf}
  \caption{AP Boxplot Summaries}
  \label{fig:ap-summaries}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{biclust-16.pdf}
  \caption{Biclustering $N = 16$}
  \label{fig:biclust-16}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{bubbleplot-16.pdf}
  \caption{Bubbleplot $N = 16$}
  \label{fig:bubbleplot-16}
\end{figure}

>>>>

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    Name & Type & Format & Description \\
    \hline
    nms\_d1 & byte & \%8.0g & cardiovascular \\
    nms\_d2 & byte & \%8.0g & sleep/fatigue \\
    nms\_d3 & byte & \%8.0g & mood/cognition \\
    nms\_d4 & byte & \%8.0g & percep/hallucinations \\
    nms\_d5 & byte & \%8.0g & attention/memory \\
    nms\_d6 & byte & \%8.0g & gastrointestinal \\
    nms\_d7 & byte & \%8.0g & urinary \\
    nms\_d8 & byte & \%8.0g & sexual function \\
    nms\_d9 & byte & \%8.0g & miscellaneous \\
    tremor & float & \%9.0g & tremor \\
    bradykin & float & \%9.0g & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & \%9.0g & rigidity \\
    axial & float & \%9.0g & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
    pigd & float & \%9.0g & postural instability and gait difficulty \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.76&  3.32&   0-24 \\
nms\_d2&   8.71&  8.76&   0-48 \\
nms\_d3&   8.70& 11.83&   0-60 \\
nms\_d4&   1.65&  3.94&   0-33 \\
nms\_d5&   5.22&  7.44&   0-36 \\
nms\_d6&   5.67&  6.92&   0-36 \\
nms\_d7&   8.02&  9.09&   0-36 \\
nms\_d8&   3.57&  5.97&   0-24 \\
nms\_d9&   6.99&  7.74&   0-48 \\
tremor&   2.59&  2.63&   0-12 \\
bradykin& 2.49&  1.39&   0-6 \\
rigidity& 2.34&  1.36&   0-6 \\
axial&    3.28&  2.75&   0-12 \\
pigd&     3.36&  2.77&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{pca-eigenvalues.png}
  \caption{Scree test: eigenvalues by factor}
  \label{fig:pca-eigenvalues}
\end{figure}

@@

\clearpage
\section{$k$-means}
\subsection{Identifying optimal number of clusters}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-wss-error.png}
  \caption{Scree test: WSS error by cluster size}
  \label{fig:kmeans-wss-error}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.png}
  \caption{Gap statistic by cluster size}
  \label{fig:gap-statistic}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{asw.png}
  \caption{Average silhouette width by cluster size}
  \label{fig:asw}
\end{figure}

@@

\subsection{Cluster statistics}
% CLUSTERS: 2
% ================================
% Sizes: 201 580
% WithinSS: 4248.585 4132.434
% Sum WithinSS: 8381.019
% CLUSTERS: 3
% ================================
% Sizes: 420 231 130
% WithinSS: 2618.368 1973.82 3076.542
% Sum WithinSS: 7668.73
% CLUSTERS: 4
% ================================
% Sizes: 61 372 145 203
% WithinSS: 1481.25 1845.389 2147.988 1609.555
% Sum WithinSS: 7084.183
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    $k$ & $n$ & Within SS & sum(Within SS) \\
    \hline
    2 & 201/580 & 4248.585/4132.434 & 8381.019 \\
    3 & 420/231/130 & 2618.368/1973.82/3076.542 &  7668.73 \\
    4 & 61/372/145/203 & 1481.25/1845.389/2147.988/1609.555 & 7084.183
  \end{tabular}
  \caption{Cluster statistics}
  \label{tab:cluster-statistics}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-2.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 2$}
  \label{fig:kmeans-silhouette-2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-3.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 3$}
  \label{fig:kmeans-silhouette-3}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-silhouette-4.pdf}
  \caption{$k$-means cluster silhouette plot, $k = 4$}
  \label{fig:kmeans-silhouette-4}
\end{figure}

\subsection{Decision trees based on clusters}
% seed = 911
% CLUSTERS: 2
% ================================
% Complexity Parameter: 0.03482587
% 10-fold CV error: 0.134443
% Root node error: 0.2573624
% CLUSTERS: 3
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.1946223
% Root node error: 0.4622279
% CLUSTERS: 4
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.2483995
% Root node error: 0.5236876
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l|l|l}
    $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    validation} & Root Feature &
    Root Error & Figure \\
    \hline
    2 & 0.0348 & 0.134 & axial $\geq$ 0.44 & 0.257 & Figure~\ref{fig:kmeans-dtree-2} \\
    3 & 0.0100 & 0.194 & bradykin $<$ 0.0041 & 0.462 & Figure~\ref{fig:kmeans-dtree-3} \\
    4 & 0.0100 & 0.248 & bradykin $<$ 0.0041 & 0.523 & Figure~\ref{fig:kmeans-dtree-4} \\
  \end{tabular}
  \caption{$k$-kmeans decision trees statistics}
  \label{tab:k-means-dtrees}
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-2.pdf}
  \caption{Decision Tree from $k$-means clustering, 2 clusters}
  \label{fig:kmeans-dtree-2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-3.pdf}
  \caption{Decision Tree from $k$-means clustering, 3 clusters}
  \label{fig:kmeans-dtree-3}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-kmeans-pruned-unscaled-4.pdf}
  \caption{Decision Tree from $k$-means clustering, 4 clusters}
  \label{fig:kmeans-dtree-4}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-2.pdf}
  \caption{Cluster Summaries, $k = 2$}
  \label{fig:kmeans-summaries-2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-3.pdf}
  \caption{Cluster Summaries, $k = 3$}
  \label{fig:kmeans-summaries-3}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-summaries-4.pdf}
  \caption{Cluster Summaries, $k = 4$}
  \label{fig:kmeans-summaries-4}
\end{figure}

\clearpage
\section{Affinity Propagation}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{ap-silhouette.pdf}
  \caption{AP silhouette plot, $k = 8$}
  \label{fig:ap-silhouette}
\end{figure}

\clearpage
\section{Hierarchical Clustering}

Four dissimilarity methods were used with a euclidean distance matrix.
Dendrograms available in Figure~\ref{fig:hc-dendrograms}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{hc-dendrograms.pdf}
  \caption{Dendrograms}
  \label{fig:hc-dendrograms}
\end{figure}

\subsection{Cutting Trees}
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    Method & Condition & n & Figure \\
    \hline
    Complete & $k = 4$ & 4 (665/74/35/7) & ~\ref{fig:hc-summaries-complete-k4} \\
    Complete & \texttt{dynamicTreeCut}\tablefootnote{Package \texttt{dynamicTreeCut} in R (Langfelder P,
  Zhang B, Horvath S (2007)). Hybrid method, minimum cluster selection
parameters} & 11 (7/270/138/83/79/49/46/39/35/35) &
    ~\ref{fig:hc-summaries-complete-dynamic} \\
    Ward & $k = 4$ & 4 (294/236/120/131) & ~\ref{fig:hc-summaries-ward-D-k4} \\
    Ward & $h = 60$ & 6 (97/236/120/197/91/40) &
    ~\ref{fig:hc-summaries-ward-D-h60} \\
  \end{tabular}
  \caption{Clusters from Tree Cutting}
  \label{tab:tree-cutting}
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{hc-summaries-complete-k4.pdf}
  \caption{Using maximum (complete linkage) dissimilarity, cutting tree for $k = 4$}
  \label{fig:hc-summaries-complete-k4}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{hc-summaries-complete-dynamic.pdf}
  \caption{Using maximum (complete linkage) dissimilarity, cutting tree
    dynamically}
  \label{fig:hc-summaries-complete-dynamic}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{hc-summaries-ward-D-k4.pdf}
  \caption{Using Ward (1963) dissimilarity, cutting tree for $k = 4$}
  \label{fig:hc-summaries-ward-D-k4}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{hc-summaries-ward-D-h60.pdf}
  \caption{Using Ward (1963) dissimilarity, cutting tree at $h = 60$}
  \label{fig:hc-summaries-ward-D-h60}
\end{figure}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{l|l}
  Cluster & Size \\
  \hline
  1 & 54 \\
  2 & 68 \\
  3 & 25 \\
  4 & 259 \\
  5 & 102 \\
  6 & 68 \\
  7 & 185 \\
  8 & 20 \\
  \end{tabular}
  \caption{AP Cluster Sizes}
  \label{tab:ap-cluster-sizes}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{ap-summaries.pdf}
  \caption{AP Boxplot Summaries}
  \label{fig:ap-summaries}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{biclust-16.pdf}
  \caption{Biclustering $N = 16$}
  \label{fig:biclust-16}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{bubbleplot-16.pdf}
  \caption{Bubbleplot $N = 16$}
  \label{fig:bubbleplot-16}
\end{figure}
